---
name: Alexander
surname: Piper
position: "PhD Candidate"
address: "AgriBio, Centre for AgriBioscience, 5 Ring Road, Bundoora Victoria"
phone: "+61 488 040 119"
profilepic: profile.JPG
#www: alexpiper.github.io/
email: "alexander.piper@agriculture.vic.gov.au"
twitter: bigsnpenergy
github: alexpiper
#linkedin: alexander-piper-702bb693/
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Alexander Piper is a graduate student with Agriculture Victoria Research and La Trobe University whose research uses high-throughput sequencing and computational biology to improve detection and control of insect pests."
output: vitae::awesomecv
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(scholar)
library(tidyverse)
```

# Education

```{r Education}
tibble::tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where, ~Why,
  "PhD Candidate", "2017-Present", "La Trobe University", "Melbourne, Australia","Thesis title: Genomic tools for detection of insect pests",
  "Bachelor of Science (Biology)", "2015", "Queensland University of Technology", "Brisbane, Australia","Biotechnology and Genetics Minor",
    "Bachelor of Science (Biology)", "2015", "Queensland University of Technology", "Brisbane, Australia","Chemistry Minor"
) %>% 
  detailed_entries(what=Degree, when=Year, with=Institution, where=Where, why=Why)
```

# Research experience

```{r}
tibble::tribble(
  ~Year, ~Position, ~Institution, ~Where, ~Why,
"2017-Present", "PhD Candidate", "La Trobe University", "Melbourne, Australia", "Developing a quantitative Metabarcoding pipeline for high-throughput identification of invasive insects within mixed trap samples",
"2017-Present", "PhD Candidate", "La Trobe University", "Melbourne, Australia", "Using population genomics to understand the historical demography and seasonal population dynamics of a pest fruit fly",
  
"2017-Present", "Research Scientist (Part time) - Computational Biology", "Agriculture Victoria Research", "Melbourne, Australia", "Analytical support for the cross-industry iMapPESTS Metabarcoding surveillance program.",
  
"2016-2017", "Research Scientist - Chemical Ecology", "Agriculture Victoria Research", "Melbourne, Australia", "Utilizing microbial symbionts of horticultural pests to develop novel insect attractants.",
  
"2015-2016", "Laboratory Assistant - Microbial Ecology", "Queensland University of Technology","Brisbane, Australia", "Isolation and physiological characterisation of environmental microbes from insects and host plants.",

"2015", "Vacation Research Scholar — Microbiology", "Queensland University of Technology","Brisbane, Australia", "Exploratory research into fungal symbionts of an agricultural pest insect funded by a university scholarship.",
  
"2015", "B.Sc. Capstone Research Project — Molecular Ecology", "Queensland University of Technology","Brisbane, Australia", "Developing an environmental DNA PCR assay for early detection of an invasive freshwater fish species.",

"2013-2014", "Undergraduate Research — Physiological Genomics","Queensland University of Technology","Brisbane, Australia", "Assisting a PhD student mapping the transcriptome of cnidarian sea anemone."
)%>% detailed_entries(with = Institution, when = Year, what=Position, where=Where, why = Why)
```

# Journal articles

```{r publications}
published <- scholar::get_publications("zrxQbOsAAAAJ&hl") %>% 
  distinct(title, .keep_all = TRUE) %>%
  transmute(bibtype = "Article", author = as.character(author),
            title = as.character(title),
            journaltitle = as.character(journal), year) 

prep <- tibble::tribble(~bibtype, ~author, ~title, ~year, ~note,
"Unpublished","J. Batovska, A.M. Piper, I. Valenzuela, J.P. Cunningham and M.J. Blacket", "Developing a non-destructive metabarcoding protocol for detection of Invasive insects in bulk trap catches", 2019,"(In Prep)"
) 

bind_rows(published, prep) %>% 
  dplyr::arrange(-year, note) %>%
  mutate(key = row_number()) %>% 
  RefManageR::as.BibEntry() %>% 
  RefManageR::WriteBib("piper.bib")


bibliography_entries("piper.bib") 
```

  
# Conference talks

```{r talks}
tibble::tribble(
  ~name, ~event, ~ location, ~year,
  "An updated molecular toolbox for Biosecurity","Agriculture Victoria Regional Science Conference","Tatura, Victoria", 2019,
 "Detecting the unexpected, DNA Metabarcoding for high-throughput insect surveillance", "AgriBio Science Conference", "Melboure, Victoria", 2018,
  "Quality control considerations for Metabarcoding", "Victorian DNA Barcoding Workshop","Melbourne, Victoria", 2018,
  "Detecting the unexpected, DNA Metabarcoding for high-throughput insect surveillance", "Australian Entomological Society Conference", "Alice Springs, Northern Territory", 2018,
  "Quality control considerations for Metabarcoding", "iMapPESTS Metabarcoding Workshop", "Melbourne, Victoria", 2018,
  "The importance of Yeasts in the ecology and control of Queensland Fruit Fly", "SciPlant 17", "Brisbane, Queensland", 2017,
  "Yeast-insect interactions in the Queensland fruit fly (Bactrocera tryoni)", "Australian Entomological Society Conference", "Terrigal, New South Wales", 2017,
  "A microbial hypothesis for Queensland fruit fly host selection", "Biology of Tephritid Fruit Flies Meeting IV", "Melbourne, Victoria", 2016
) %>%  detailed_entries(
  what = name,
  when = year,
  with = event,
  where = location
  )
```

# Synergistic activities

* Member of The Society for Molecular Biology and Evolution
* Member of The Australian Bioinformatics and Computational Biology Society
* Member of The Australain Entomological Society
* Participant in the Insect Genetic Technologies Research Coordination Network (IGTRCN) 
* Reviewer for Journal of Economic Entomology

# References

* __Assoc Prof. Paul Cunningham__  
  Research Leader — Invertebrate and Weed Sciences   
  Agriculture Victoria Research  
  Phone: +613 9032 7382  
  Email: paul.cunningham@agriculture.vic.gov.au  

* __Dr. Noel Cogan__  
  Senior Research Scientist  
  Agriculture Victoria Research  
  Phone: +613 9032 7096  
  Email: noel.cogan@agriculture.vic.gov.au  
