---
surname: Alexander Piper
position: "Research Scientist"
address: "AgriBio, Centre for AgriBioscience, 5 Ring Road, Bundoora, Victoria, Australia"
phone: "+61 488 040 119"
#profilepic: profile.JPG
#www: alexpiper.github.io/
email: "alexander.piper@agriculture.vic.gov.au"
twitter: bigsnpenergy
github: alexpiper
#linkedin: alexander-piper-702bb693/
date: "`r format(Sys.time(), '%B %Y')`"
aboutme: "Alexander Piper is a Research Scientist with Agriculture Victoria Research whose research uses high-throughput sequencing and computational biology to improve detection and control of insect pests."
headcolor: 414141
output: vitae::awesomecv
sansfont: Lora
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
library(scholar)
library(tidyverse)
```

# Education

```{r Education}
tibble::tribble(
  ~ Degree, ~ Year, ~ Institution, ~ Where, ~Why,
  "Doctor of Philosophy", "2021", "School of Applied Systems Biology, College of Science, Health, and Engineering, La Trobe University", "Melbourne, Australia","Thesis title: Genomic Bio-surveillance for Insect Pests",
  "Bachelor of Science (Biology)", "2015", "Queensland University of Technology", "Brisbane, Australia","Biotechnology and Genetics Minor",
    "Bachelor of Science (Biology)", "2015", "Queensland University of Technology", "Brisbane, Australia","Chemistry Minor"
) %>% 
  detailed_entries(what=Degree, when=Year, with=Institution, where=Where, why=Why)
```

# Research experience

```{r Experience}
tibble::tribble(
  ~Year, ~Position, ~Institution, ~Where, ~Why,
"2017-2021", "Research Scientist", "Agriculture Victoria Research", "Melbourne, Australia", "Developing high-throughput sequencing assays and bioinformatic pipelines for diagnostics of insect pests.",
  
"2016-2017", "Research Scientist", "Agriculture Victoria Research", "Melbourne, Australia", "Investigating the role of microbial volatile organic compounds in the ecology of Bactrocera fruit flies, and their potential for application in novel insect attractant formulations.",
  
"2015-2016", "Laboratory Assistant", "Queensland University of Technology","Brisbane, Australia", "Isolation and physiological characterisation of environmental microbes from insects and host plants.",

"2015", "Vacation Research Scholar", "Queensland University of Technology","Brisbane, Australia", "Exploratory research into fungal symbionts of an agricultural pest insect funded by a university scholarship.",
  
"2015", "B.Sc. Capstone Research Project", "Queensland University of Technology","Brisbane, Australia", "Developing an environmental DNA PCR assay for detection of an invasive freshwater fish species.",

"2013-2014", "Undergraduate Research","Queensland University of Technology","Brisbane, Australia", "Validation of transcriptomics differentially expressed genes using qRT-PCR."
)%>% detailed_entries(with = Institution, when = Year, what=Position, where=Where, why = Why)
```

# Journal articles

```{r publications}
published <- scholar::get_publications("zrxQbOsAAAAJ&hl") %>% 
  distinct(title, .keep_all = TRUE) %>%
  transmute(bibtype = "Article", author = as.character(author),
            title = as.character(title),
            journaltitle = as.character(journal), year) %>%
  filter(!journaltitle == "bioRxiv")

prep <- tibble::tribble(~bibtype, ~author, ~title, ~year, ~notes, ~journaltitle,
"Unpublished","AM Piper, JP Cunningham, NOI Cogan, and MJ Blacket ", "DNA Metabarcoding Enables High-Throughput Detection of Spotted Wing Drosophila (Drosophila suzukii) within Unsorted Trap Catches", 2021, "Fronteirs in Ecology and Evolution (In Review)", "Fronteirs in Ecology and Evolution (In Review)"
) 

bind_rows(published, prep) %>%
  dplyr::arrange(match(bibtype, c("Unpublished","Article")), desc(year)) %>%
  mutate(key = row_number()) %>% 
  RefManageR::as.BibEntry() %>% 
  RefManageR::WriteBib("piper.bib")


bibliography_entries("piper.bib") 
```


# Preprint articles

```{r preprints}
preprint <- scholar::get_publications("zrxQbOsAAAAJ&hl") %>% 
  distinct(title, .keep_all = TRUE) %>%
  transmute(bibtype = "Article", author = as.character(author),
            title = as.character(title),
            journaltitle = as.character(journal), year)  %>%
  filter(journaltitle == "bioRxiv")


preprint %>% 
  dplyr::arrange(-year) %>%
  mutate(key = row_number()) %>% 
  RefManageR::as.BibEntry() %>% 
  RefManageR::WriteBib("piper_preprint.bib")


bibliography_entries("piper_preprint.bib") 
```

  
# Selected scientific presentations

```{r talks}
tibble::tribble(
  ~name, ~event, ~ location, ~year,
  "A DNA metabarcoding assay for early detection of Spotted Wing Drosophila within unsorted bulk trap catches.","Australian Entomological Society Conference", "Online", 2021,
  "Detecting the Unexpected: Invasive Insect Surveillance using Non-Destructive DNA Metabarcoding.","International Congress of Entomology", "Helsinki, Finland (Cancelled due to COVID19)", 2020,
  "Towards quantitative and high-throughput insect surveillance using DNA Metabarcoding.","Australian Entomological Society Conference", "Brisbane, Queensland", 2019,
  "An updated molecular toolbox for Biosecurity.","Agriculture Victoria Regional Science Conference","Tatura, Victoria", 2019,
 "Detecting the unexpected, DNA metabarcoding for high-throughput insect surveillance.", "AgriBio Science Conference", "Melboure, Victoria", 2018,
  "Quality control considerations for metabarcoding.", "Victorian DNA Barcoding Workshop","Melbourne, Victoria", 2018,
  "Detecting the unexpected, DNA metabarcoding for high-throughput insect surveillance.", "Australian Entomological Society Conference", "Alice Springs, Northern Territory", 2018,
  "Quality control considerations for metabarcoding.", "iMapPESTS metabarcoding Workshop", "Melbourne, Victoria", 2018,
  "The importance of Yeasts in the ecology and control of the Queensland Fruit Fly.", "SciPlant 17", "Brisbane, Queensland", 2017,
  "Yeast-insect interactions in the Queensland fruit fly (Bactrocera tryoni).", "Australian Entomological Society Conference", "Terrigal, New South Wales", 2017,
  "A microbial hypothesis for Queensland fruit fly host selection.", "Biology of Tephritid Fruit Flies Meeting IV", "Melbourne, Victoria", 2016
) %>%  detailed_entries(
  what = name,
  when = year,
  with = event,
  where = location
  )
```

# Software development


```{r r-pkgs}
tribble(
  ~ pkg, ~ when, ~ role, ~ details,
  "taxreturn", "2021", "Lead developer", "An R package for curating public DNA sequence databases for metabarcoding studies.",
  "piperline", "2021", "Lead developer", "A nextflow-based metabarcoding pipeline for detection of regulated species."
) %>% 
  brief_entries(glue::glue("\\textbf{<pkg>}: <role> \\newline <details>", .open = "<", .close = ">"), paste0(when, "\\newline"), .protect = FALSE)
```

# Society memberships

* Member, Australian Bioinformatics and Computational Biology Society.

* Member, Australian Entomological Society.

* Member, Society for Molecular Biology and Evolution.

* Member, International Society for Computational Biology

# Synergistic activities

* Reviewer for VALITEST Work Package 2 - Guidelines for validation and application of non-targeted HTS diagnostic procedures in plant pest diagnostics.

* Reviewer for Molecular Ecology Resources, Evolutionary Ecology, and Journal of Economic Entomology.

* Participant in the Insect Genetic Technologies Research Coordination Network (IGTRCN).

# References

* __Assoc Prof. Paul Cunningham__  
  Research Leader — Invertebrate and Weed Sciences   
  Agriculture Victoria Research  
  Phone: +613 9032 7382  
  Email: paul.cunningham@agriculture.vic.gov.au  

* __Dr. Noel Cogan__  
  Research Leader — Molecular Genetics 
  Agriculture Victoria Research  
  Phone: +613 9032 7096  
  Email: noel.cogan@agriculture.vic.gov.au  

* __Dr. Mark Blacket__  
  Senior Research Scientist — Invertebrate and Weed Sciences   
  Agriculture Victoria Research  
  Phone: +613 9032 7333 
  Email: mark.blacket@agriculture.vic.gov.au 